### 2010-01-05 trunk/latex/paper1/contents.tex
Agora sim, cabendo em 4 páginas.

Although many software clustering algorithms have been proposed, there is
little empirical evaluation regarding the external quality of the
decompositions they find. One important test for a software clustering
algorithm is to apply it to a collection of software systems with known
reference decompositions made by experienced developers \cite{Anquetil1999}. It
is expected that a good algorithm finds decompositions that are similar to
reference decompositions, which can be measured by a similarity metric such as
MoJo or EdgeSim \cite{Mitchell2001}.

@@

Scale-free network models are algorithms that can be proven, either formally or
empirically, to synthesize networks that are scale-free. Although many
scale-free network models have been proposed, we were able to find in the
literature only two network models with built-in modules: CGW \cite{Chen2008}
and LFR \cite{Lancichinetti2009}. Also, we adapted a simple and flexible model,
BCR \cite{Bollobas2003}, to produce modules. We called the new model BCR+.

@@

In a recent work, Milo et al. \cite{Milo2004} proposed to characterize networks
by analyzing their triad concentration. A triad is a network with three
vertices in which all vertices are connected. There are only 13 distinct
triads, one for each possible configuration of directed edges, as shown in
Figure \ref{fig:triads}.

@@

Using the data set we can define S-score, a metric that represents how much a
particular network resemble software networks. It is defined as the average
similarity between the network and a sample of software networks:

@@

We used the tool igraph (\url{http://igraph.sf.net/}) to extract the TCP for
each network in the data set. Then, we measured the S-score for each software
network, which ranged from 0.83 to 0.98. The average S-score was 0.97 and the
standard deviation, 0.03. The high average S-score and the low standard
deviation show that the metric successfully characterizes software networks by
capturing their common structural patterns.

@@

\subsection{A Network Classification Model} \label{sec:classmodel}
% TODO: shorten this subsection
Although the S-score of a network tells how close it is from software networks,
it does not tell whether a network is close enough that it can be considered
software-like. What is needed is a binary classification model that
distinguishes software-like networks from the other networks. The distinction
can be made by choosing a suitable S-score threshold. Networks with S-score
below the threshold are considered dissimilar from software networks; only
networks with S-score above the threshold are considered software-like.

@@

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. The recall in this case,
though, would be too low, because most software networks would be
misclassified. So we chose the value 0.88, that is immediately above the second
greater S-score for a non-software network. With this value, we have both high
recall (95.4\%) and high precision (96.9\%).

@@

Because the possible combinations of parameter values are infinite, we have set
the number of vertices to 1000 and then varied the remaining parameters in
discrete steps. The number of modules varied between 2, 4, 8, 16, and 32. For
each of the remaining model-specific parameters, at least 3 values were chosen.
These values were chosen either to cover the domain for the parameter or, when
the domain was infinite, were tuned to approximate characteristics networks in
the software network data set. In total, 9,500 networks were generated by the
BCR+ model, 38,790 by the CGW model, and 1,296 by the LFR model.

@@

Each synthesized network was classified as software-like or non software-like,
using the classification model presented in section \ref{sec:classmodel}. The
results are summarized in Table \ref{tab:results}.

@@

The use of synthetic data is common in distributed computing research, but
still under-explored in software engineering research. Because many reverse
engineering tasks rely on dependency data, we expect this
work to have impact beyond the software clustering community.

>>>>

Although many software clustering algorithms have been proposed, there is
little empirical evaluation regarding the external quality of the
decompositions they find, i.e., how close the decompositions are from reference
decompositions made by experienced developers \cite{Anquetil1999}. It is
expected that a good algorithm finds decompositions that are similar to
reference decompositions, which can be measured by a similarity metric such as
MoJo or EdgeSim \cite{Mitchell2001}.

@@

Scale-free network models are algorithms that can be proven, either formally or
empirically, to synthesize networks that are scale-free. Although many
scale-free network models have been proposed, we were able to find in the
literature only two models that generate networks with built-in modules: CGW
\cite{Chen2008} and LFR \cite{Lancichinetti2009}. Also, we adapted a simple and
flexible model, BCR \cite{Bollobas2003}, to produce modules. We called the new
model BCR+.

@@

In a recent work, Milo et al. \cite{Milo2004} characterized networks by
analyzing their triad concentration. A triad is a network with three vertices
in which all vertices are connected. There are only 13 distinct triads, one for
each possible configuration of directed edges, as shown in Figure
\ref{fig:triads}.

@@

Using the data set we can define S-score, a metric that represents how much a
particular network resembles software networks. It is defined as the average
similarity between the network and a sample of software networks:

@@

We used the tool igraph (\url{http://igraph.sf.net/}) to extract the TCP for
each network in the data set. Then, we measured the S-score for each software
network, which ranged from 0.83 to 0.98, with average 0.97 and standard
deviation 0.03. The high average S-score and the low standard deviation show
that the metric successfully characterizes software networks by capturing their
common structural patterns.

@@

\subsection{A Network Classification Model} \label{sec:classmodel} % TODO:
shorten this subsection Although the S-score of a network tells how close it is
from software networks, it does not tell whether a network is close enough that
it can be considered software-like. What is needed is a binary classification
model that distinguishes software-like networks from the other networks. The
distinction can be made by choosing a suitable S-score threshold. If a network
has a S-score below the threshold, it is considered dissimilar from software
networks; otherwise, it is considered software-like.

@@

To get 100\% precision, the threshold needs to be 0.98, so the non-software
network with highest S-score is below the threshold. But the recall in this
case would be low, because most software networks would be misclassified, so we
chose the value 0.88, that is immediately above the second greater S-score for
a non-software network. With this value, we have both high recall (95.4\%) and
high precision (96.9\%).

@@

Because the possible combinations of parameter values are infinite, we have set
the number of vertices to 1000 and then varied the remaining parameters in
discrete steps. The number of modules varied between 2, 4, 8, 16, and 32. For
each of the remaining model-specific parameters, at least 3 values were chosen.
When possible, the values were chosen to cover the entire parameter domain. In
the case of unbound parameters, the values were chosen to approximate
characteristics of the networks in the software network data set. In total,
9,500 networks were generated with the BCR+ model, 38,790 with the CGW model,
and 1,296 with the LFR model.

@@

Each synthesized network was classified as software-like or non software-like,
using the classification model presented in Section \ref{sec:classmodel}. The
results are summarized in Table \ref{tab:results}.

@@

The use of synthetic data is common in distributed computing research, but
still underexplored in software engineering research. Because many reverse
engineering tasks rely on dependency data, we expect this
work to have impact beyond the software clustering community.
